---
title: "Make a map of focal trees"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/2021-05-28_create-map/")
```

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("sp")
library("ggmap")
library("geojsonio")
library("leaflet")
library("osmdata")
```

## Read in gpx file and clean up
This file has come straight from the garmin gps

```{r read-gpx, eval = FALSE}
plotKML::readGPX(here::here("data", "raw", "waypoints.gpx")) %>%
  map_df(~.) %>%
  rename(id = name) %>%
  mutate(taxa = case_when(sym == "Park" ~ "Prunus_spinosa",
                          sym == "Flag, Blue" ~ "Crataegus_monogyna")) %>%
  select(-sym) %>%
  arrange(id) -> waypoints
```


## Or use co-ordinates from csv

```{r read-csv}
read.csv(here::here("data", "raw", "tree_id_checklist.csv")) %>%
  drop_na(long_dd, lat_dd) -> tree_list

```


## Make a static map

```{r ggmap, message=FALSE}
bbox <- make_bbox(c(min(tree_list$long_dd) - 0.001, 
                    max(tree_list$long_dd)) + 0.001, 
                  c(min(tree_list$lat_dd) - 0.001, 
                    max(tree_list$lat_dd) + 0.001))

osmdata::opq(bbox = bbox) %>%
  add_osm_feature(key = "highway") %>%
  osmdata_sf() -> osm_trails

tree_list %>%
  filter(focal == TRUE) %>%
  filter(tree_id != "23" & tree_id != "26") -> tree_list_focal

get_map(bbox, source = "stamen", force = TRUE, maptype = "terrain") %>%
  ggmap() +
  geom_sf(data = osm_trails$osm_lines,
          inherit.aes = FALSE, colour = "grey60") +
  geom_point(data = tree_list_focal,
             colour = "red", shape = 4,
           aes(long_dd, lat_dd)) +
  geom_text(data = tree_list_focal,
             aes(long_dd, lat_dd, label = tree_id), hjust = -0.5) +
  theme_void()

```

## Make an interactive map

### make the data a SpatialPointsDataFrame object 

```{r make-spdf}
SpatialPointsDataFrame(coords = select(tree_list, long_dd, lat_dd), 
                       data = tree_list, 
                       proj4string = CRS("+init=epsg:4326")) -> tree_list_sp
```

### make a map with leaflet
This is interactive in RStudio and if we knit to html, but GitHub won't render it.

```{r leaflet-map}
# colours
pal <- colorFactor(c("gold", "navy"), c("FALSE", "TRUE"))

leaflet(data = tree_list_sp)%>% 
  addCircleMarkers(label = ~tree_id, color = ~pal(focal), 
                            fillOpacity = 0.8, stroke = FALSE, radius = 5) %>% 
  addTiles() %>%
  addLegend(pal = pal, values = ~focal)
```

### make a geoJSON file
We can export the SpatialPointsDataFrame to a geoJSON, which GitHub will render as an interactive map, annotated
with our geodata. This gives us a browsable, online version we can refer to.

```{r geojson}
# convert to geojson
tree_list_geojson <- geojson_json(tree_list_sp)

# write
geojson_write(tree_list_geojson, file = "figures/2021-05-28_create-map/focal-trees-map.geojson")
```

__view the interactive map online__ [__here__](figures/2021-05-28_create-map/focal-trees-map.geojson)
