---
title: "Make a map of focal trees"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/2021-05-28_create-map/")
```

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("sp")
library("ggmap")
library("geojsonio")
library("leaflet")
```

## Read in gpx file and clean up
This file has come straight from the garmin gps

```{r read-gpx}
plotKML::readGPX(here::here("data", "raw", "waypoints.gpx")) %>%
  map_df(~.) %>%
  rename(id = name) %>%
  mutate(taxa = case_when(sym == "Park" ~ "Prunus_spinosa",
                          sym == "Flag, Blue" ~ "Crataegus_laevigata")) %>%
  select(-sym) -> waypoints
```

## Make a static map

```{r ggmap, message=FALSE}
bbox <- make_bbox(c(min(waypoints$lon), max(waypoints$lon)), 
                  c(min(waypoints$lat), max(waypoints$lat)))

get_map(bbox, source = "stamen", force=TRUE) %>%
  ggmap() +
  geom_point(data = waypoints, 
           aes(lon, lat, colour = taxa))
```

It's not very detailed - we can't see any of the trails

## Make an interactive map

### make the data a SpatialPointsDataFrame object 

```{r make-spdf}
SpatialPointsDataFrame(coords = select(waypoints, lon, lat), 
                       data = waypoints, 
                       proj4string = CRS("+init=epsg:4326")) -> waypoints_sp
```

### make a map with leaflet
This is interactive in RStudio and if we knit to html, but GitHub won't render it.

```{r leaflet-map}
# colours
pal <- colorFactor(c("navy", "red"), c("Prunus_spinosa", "Crataegus_laevigata"))

leaflet(data = waypoints_sp)%>% 
  addCircleMarkers(label = ~id, color = ~pal(taxa), 
                            fillOpacity = 0.8, stroke = FALSE, radius = 5) %>% 
  addTiles() %>%
  addLegend(pal = pal, values = ~taxa)
```

### make a geoJSON file
We can export the SpatialPointsDataFrame to a geoJSON, which GitHub will render as an interactive map, annotated
with our geodata (taxa, id). This gives us a browsable, online version we can refer to.

```{r geojson}
# convert to geojson
waypoints_geojson <- geojson_json(waypoints_sp)

# write
geojson_write(waypoints_geojson, file = "figures/2021-05-28_create-map/focal-trees-map.geojson")
```

[__View map here!__](figures/2021-05-28_create-map/focal-trees-map.geojson)
