---
title: "First look at mapping data; cleaning and connectivity"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/2022-07-07_explore-mapping-data/")
source(here::here("code", "secrets.R"))
library("emo")
```

We want to conduct herbivory surveys on individuals at high and low conspecific densities. To identify them we'll need a measure of connectivity. I'm going to take a first look at the mapping data that has been collected so far and have a go at calculating connectivity for all of the trees.

```{r packages}
library("tidyverse")
library("janitor")
library("googledrive")
library("here")
library("rdist")
```

## Cleaning

```{r clean}
googledrive::drive_auth(email = "eleanor.elizabeth.j@gmail.com",
                        scopes = "https://www.googleapis.com/auth/drive") 

files <- drive_ls(GDRIVE_FOLDER_URL,
                  pattern = "^[0-9]{2}.csv") # don't use incomplete plots

files  %>%
  mutate(id = sprintf("https://docs.google.com/uc?id=%s&export=download", id)) %>%
  pull(id) -> urls

names(urls) <- pull(files, name)

data_list <- lapply(urls, read.csv, na.strings = c("","NA"))

repair_col_names <- function (plot) {
  plot %>%
    rename_with(recode, "BHD" = "DBH") %>%
    rename_with(recode, "BHC" = "CBH") %>%
    rowwise() %>%
    mutate(DBH = ifelse("DBH" %in% names(.), DBH, NA),
           CBH = ifelse("CBH" %in% names(.), CBH, NA),
           Other.HC = ifelse("Other.HC" %in% names(.), Other.HC, NA)) %>% 
    ungroup() %>%
    mutate(across(c(DBH, CBH, Other.HC, Other.HD), as.numeric)) %>%
    mutate(Notes = as.character(Notes)) %>%
    select(-contains("X"))
}

data_list_col_repair <- lapply(data_list, repair_col_names)

repair_tree_names <- function (plot, plot_name) {
  plot %>% 
    # remove any letters
    mutate(Name = gsub("[a-zA-Z]+", "", x = Name, ignore.case = TRUE)) %>%
    #remove leading zeros
    mutate(Name = gsub("^0+", "", Name) ) %>% 
    # rename 1st row as focal tree
    mutate(Name = replace(Name, pluck(1, 1), 
                          paste0("focal_", readr::parse_number(plot_name)))) %>%
    mutate(Name = as.factor(Name))
}

data_list_name_repair <- imap(data_list_col_repair, repair_tree_names)
      
all_plots <- bind_rows(data_list_name_repair, .id = "plot")

all_plots %>%
  clean_names() %>% 
  fill(c("name", "longitude", "latitude", "solution_status", 
         "cs_name", "form", "reproductive")) %>%
  mutate(plot = gsub(".csv", "", x = plot)) %>%
  rename(tree_id = name) %>%
  group_by(plot) %>%
  mutate(stem_id = 1:n()) %>%
  ungroup() %>% 
  add_count(tree_id, plot, name = "n_stems") %>% 
  rowwise() %>%
  mutate(dbh = ifelse(is.na(dbh), cbh/pi, dbh)) %>%
  ungroup() %>%
  mutate(status = ifelse(is.na(status), "A", status)) %>%
  mutate_if(is.character, tolower) %>%
  mutate(tree_id = paste("tree", tree_id, sep = "_"),
         stem_id = paste("stem", stem_id, sep = "_")) -> clean_plots

```

## Calculate distances

For the ForestGeo data they use the diameter at breast height (`dbh`) of the largest stem of a multi-stemmed tree - so I have done that here too.

```{r}
clean_plots %>%
  drop_na(dbh) %>% # this will drop dbh that were below breast height
  filter(status != "D") %>% 
  group_by(plot, tree_id) %>% 
  slice_max(dbh, with_ties = FALSE) %>% 
  ungroup() -> largest_stems

calculate_dist <- function (plot_id, data) {
  
  filter(data, plot == plot_id) -> plot_data

	plot_data %>%
    select(latitude, longitude) -> plot_matrix

	rdist::pdist(plot_matrix[,c("latitude", "longitude")], 
	             metric = "euclidean") -> dists
	
	as.data.frame(dists) -> dists_df

	unlist(plot_data$tree_id) -> colnames(dists_df) 

	cbind(plot_data, dists_df)
	
}

largest_stems %>% 
  distinct(plot) %>%
  pull(plot) -> plot_id_list

distance_dfs <- lapply(plot_id_list, calculate_dist, 
                       data = largest_stems)
```

## Calculate connectivity

Connectivity is proving harder to calculate than I thought. Compared to last time, when we calculated connectivity for one trap based on trees in a radius, we're now trying to calculate connectivity values for *every* tree in the plot.. `r emo::ji("cold_sweat")`

I might try and think of another way to identify individuals at high and low conspecific densities. This seems like too much effort and computing power for this task.

Think I'll just stick to getting the connectivity of each of the focal trees for now - just to see what it looks like. 

```{r}
all_distance_dfs <- dplyr::bind_rows(distance_dfs)

all_distance_dfs %>%
  filter(grepl("focal", tree_id)) %>%
  pull(tree_id) -> focal_tree_id_list

calculate_connectivity <- function (data, focal_tree_id) {
  data %>%
    group_by(plot, tree_id) %>%
    mutate(x = dbh * exp((-0.02) * eval(parse(text = focal_tree_id)) ) ) %>%
    ungroup() %>%
    group_by(plot) %>%
    summarise(tree_id = paste(focal_tree_id),
              connectivity = sum(x), .groups = "drop")
}

connectivity_dfs <- lapply(focal_tree_id_list, 
                              calculate_connectivity, data = all_distance_dfs)

connectivity_dfs %>%
  lapply(drop_na, connectivity) %>%
  dplyr::bind_rows() -> all_connectivity_dfs

knitr::kable(all_connectivity_dfs)

```

We are missing connectivity values for plots `22`, `33` and `35` - this is because they have no recorded `dbh` and thus they were filtered out when I created `largest_stems`. Not sure why they have no `dbh`, will check with Matt. Think I have this data recorded somewhere anyway.

## Other Hawthorn abundance metrics

Let's do some sanity checks. Does connectivity roughly align with the number of trees per plot?

Looking at `n_trees` per plot will also help us to get an idea about the range of Hawthorn densities in the mapped plots so far.

```{r n-trees}
all_distance_dfs %>%
  group_by(plot) %>%
  summarise(n_trees = n_distinct(tree_id)) -> sum_trees

check_data <- left_join(all_connectivity_dfs, sum_trees, by = "plot")

check_data %>%
  pivot_longer(cols = c("connectivity", "n_trees")) %>%
  ggplot(aes(
    x = plot,
    y = value,
    group = name,
    colour = name
  )) +
  geom_path() +
  theme_bw() +
  facet_grid(rows = "name", scales = "free")
      
```

Kind of... something weird going on with plot 24. Perhaps lots of small trees, or many trees which are far from the focal tree? I guess they are measuring quite different things.

Let's also calculate the summed basal area to compare.

```{r basal-area}
largest_stems %>%
  rowwise() %>%
  mutate(basal_area = ((dbh / 2)**2) * pi ) %>%
  ungroup() %>%
  group_by(plot) %>%
  summarise(sum_basal_area = sum(basal_area), .groups = "drop") -> sum_area

check_data_2 <- left_join(check_data, sum_area, by = "plot")

check_data_2 %>%
  pivot_longer(cols = c("connectivity", "n_trees", "sum_basal_area")) %>%
  ggplot(aes(
    x = plot,
    y = value,
    group = name,
    colour = name
  )) +
  geom_path() +
  theme_bw() +
  facet_grid(rows = "name", scales = "free")
```

Summed basal area matches connectivity much more than `n_trees` - which is good and as expected since `dbh` was used to calculate both of those values.

Before we start using these connectivity values properly we are going to have to think about edge effects, i.e. there could be loads of Hawthorn just outside of the plot boundaries.

Anyway, I'm glad that connectivity seems to align with these other abundance metrics and there seems to be a bit of variation in Hawthorn abundance between the plots so far. Most have low abundance but I think the most difficult (Hawthorn abundant) plots have been saved to last. So we may yet see more Hawthorn dense plot data.
