---
title: "Test for spatial autocorrelation"
author: "Eleanor Jackson"
date: '09 February, 2023'
always_allow_html: true
output: 
  github_document:
    keep_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/2023-02-09_test-spatial-autocorrelation/")
```

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
library("sf")
library("ncf")
library("spdep")
library("sp")
  
```

```{r data}
readRDS(here::here("data", "clean", "hawthorn_plots.rds")) %>%
  select(plot, tree_id, longitude, latitude, dbh, reproductive) %>%
  filter(tree_id == "tree_0") %>%
  mutate(tree_id = as.numeric(plot)) %>% 
  left_join(readRDS(here::here("data", "raw", "fruit_drop_data.rds")), 
            by = "tree_id") %>%
  sf::st_as_sf(coords = c("latitude", "longitude"), 
               crs = 4326, remove = FALSE) %>%
  distinct(tree_id, .keep_all = TRUE) %>%
  drop_na(proportion_dropped) -> focal_trees

# Missing mapping data for trees 29 and 23,
# 34 and 35 were only added for the pollination work,
# so n = 30

```

## Correlogram

A correlogram shows the correlation coefficient for the series lagged (in distance) by one delay at a time. e.g. at lag one you’re looking at the correlation between a point and it’s nearest neighbour. At lag two you’re looking at the correlation between a point and it’s second nearest neighbour.

Values close to 1 indicate clustering while values close to -1 indicate dispersion. A random arrangement would give a value that is close to 0.

```{r correlogram}
# using 10 m increments
plot(ncf::correlog(x = focal_trees$latitude, y = focal_trees$longitude, 
                   z =  focal_trees$proportion_dropped,
                   latlon = TRUE, increment = 0.01))

```

Here, values significant at a nominal (two-sided) 5%-level are represented by filled circles and non-significant values by open circles. 


## Moran's I

The expected value of Moran's I under the null hypothesis of no spatial autocorrelation is `-1/(n-1)`, which equals `-0.03448276` for this data. 

Values significantly below `-1/(n-1)` indicate negative spatial autocorrelation and values significantly above `-1/(n-1)` indicate positive spatial autocorrelation. 

```{r morans-i}
# detect nearest neighbours within a 15 km radius (i.e. whole island)
spdep::dnearneigh(focal_trees$geometry, 0, 15, 
                  row.names = focal_trees$tree_id,  
                  longlat = TRUE) -> nn

# add spatial weights
lw <- spdep::nb2listw(nn, style = "W", zero.policy = TRUE) 

# Moran's test for spatial autocorrelation
spdep::moran.test(focal_trees$proportion_dropped, lw) 

```

Moran's I is below `-1/(n-1)`, indicating negative spatial correlation, but it's not a significant result.

## Variogram

A variogram describes the variance of the difference between values at two locations. 

Here, the x axis represents the distance between traps (km) and each point represents a pair of observations. The distance at which the variogram stops increasing and flattens out is the range. Traps which are closer together than this distance are spatially autocorrelated.

```{r variogram}
sp::coordinates(focal_trees) = ~latitude+longitude

plot(
    gstat::variogram(proportion_dropped ~ 1, data = focal_trees, 
        width = 0.001, cutoff = 4000) 
)

```

Looks to me like we don't have any spatial autocorrelation - the semivariance is fairly flat and quite random.
