---
title: "Survival analysis for hawthorn dispersal"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

```{r packages, message=FALSE, warning=FALSE}
library("ggsurvfit")
library("tidyverse")
library("here")
library("lubridate")
library("brms")
library("bayestestR")
library("rstanarm")
library("survival")
library("coxme")
```

Need to reformat data into typical "survival data" format with one row for each fruit and time point which indicates whether it was on the tree`0` or dropped `1`.

```{r}

fruit_counts_21_22 <- 
  read.csv(here::here("data", "raw", "fruit_drop_data.csv"), 
           header = TRUE, na.strings = c("", "NA"))

fruit_counts_21_22 %>%
  mutate(date = dmy(date), branch = str_sub(branch_id, -1)) %>%
  mutate(branch = tolower(branch)) %>%
  mutate(branch_id = paste0(tree_id, branch)) %>%
  mutate(survey = case_when(
    date == "2022-03-04" | date == "2022-03-06" | date == "2022-03-07" ~ 8,
    date == "2022-01-28" | date == "2022-01-27" ~ 7,
    date == "2021-12-21" | date == "2021-12-20" ~ 6,
    date == "2021-11-25" | date == "2021-11-24" ~ 5,
    date == "2021-10-29" | date == "2021-10-28" ~ 4,
    date == "2021-09-29" | date == "2021-09-28" | date == "2021-09-27" ~ 3,
    date == "2021-09-02" ~ 2,
    date == "2021-08-06" | date == "2021-08-05" ~ 1,
  )) %>%
  mutate(exclusion = ifelse(branch == "d" | branch == "e" | branch == "f",
                            FALSE, TRUE)) %>%
  group_by(branch_id) %>%
  reframe(across(everything()),
            length_cm = median(length_cm, na.rm = TRUE)) %>%
  select(- notes) -> survey_data_21_22

# exclusion cages were added on survey 3 and counted on these
# branches for the first time

```

```{r}
survey_data_21_22 %>% 
  filter(exclusion == FALSE) %>% 
  filter(survey >2) %>% 
  group_by(branch_id) %>% 
  mutate(total_fruit = max(n_fruit)) %>% 
  ungroup() -> survival_data
```

```{r}
survival_data %>% 
  ggplot(aes(y = n_fruit/total_fruit, 
           x = survey, 
           group = branch_id, 
           colour = branch_id)) +
  geom_point() +
  geom_path() +
  facet_wrap(~tree_id) +
  theme(legend.position = "none")


```

```{r}
survival_data %>% 
  group_by(branch_id) %>% 
  mutate(days = as.numeric(
    difftime(date, 
             "2021-09-27", 
             unit = "day")
    ) ) %>% 
  ggplot(aes(y = n_fruit/total_fruit, 
           x = days, 
           group = branch_id, 
           colour = branch_id)) +
  geom_point() +
  geom_path() +
  facet_wrap(~tree_id) +
  theme(legend.position = "none")
```

```{r}
survival_data %>% 
  mutate(n_dropped = total_fruit - n_fruit) %>% 
  uncount(n_dropped) %>% 
  mutate(status = 1) -> dead_counts

survival_data %>% 
  uncount(n_fruit) %>% 
  mutate(status = 0) -> alive_counts

bind_rows(dead_counts, alive_counts) -> long_data
```


```{r}
readRDS(here::here("data", "clean", "fruit_dispersal.rds")) %>%
  select(branch_id, 
         repro_connectivity, 
         non_repro_connectivity, 
         dbh) %>% 
  right_join(long_data, by = "branch_id") %>% 
  mutate(repro_connectivity_sc = scale(repro_connectivity),
         non_repro_connectivity_sc = scale(non_repro_connectivity),
         dbh_sc = scale(dbh),
         tree_id = as.factor(tree_id),
         exclusion = as.factor(exclusion)
  ) %>% 
  mutate(days = as.numeric(
    difftime(date, 
             "2021-09-27", 
             unit = "day")
    ) ) %>% 
  mutate(days = round(days),
         survey = survey -3) -> survival_data_sc
```

# survival package

following along from here: https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html

```{r}
Surv(survival_data_sc$days, survival_data_sc$status)[1:10]
```

```{r}
s1 <- survfit(Surv(days, status) ~ 1, data = survival_data_sc)
glimpse(s1)
```
## Kaplan-Meier plots

```{r}
ggsurvfit::survfit2(Surv(days, status) ~ 1, data = survival_data_sc) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval()
```

The confidence interval is very narrow and hard to see.

Can we tease anything out of this? Try colouring by tree id:

```{r}
pal_29<-viridisLite::viridis(n = 29, option = "mako")

ggsurvfit::survfit2(
  Surv(days, status) ~ 
    repro_connectivity_sc, data = survival_data_sc) %>% 
  ggsurvfit() +
  scale_color_manual(values = pal_29) +
  scale_fill_manual(values = pal_29) +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  #add_confidence_interval() +
  theme(legend.position = "none") 

```

Maybe we can turn the predictor variables into categories (high, medium, low) to look for patterns in how they might influence the relationship between survival & time.

### reproductive conspecific density

```{r}
ggplot(survival_data_sc, aes(repro_connectivity_sc)) + 
  geom_histogram(bins=20)

data_repro_split <-survival_data_sc %>% 
  mutate(conect_cat = cut(repro_connectivity_sc, breaks= 3)) 

ggsurvfit::survfit2(Surv(days, status) ~ conect_cat,
                    data = data_repro_split) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  scale_colour_viridis_d() +
  scale_fill_viridis_d()
```

### non-reproductive conspecific density

```{r}
ggplot(survival_data_sc, aes(non_repro_connectivity_sc)) + 
  geom_histogram(bins=20)

data_repro_split <-survival_data_sc %>% 
  mutate(conect_cat = cut(non_repro_connectivity_sc, breaks= 3)) 

ggsurvfit::survfit2(Surv(days, status) ~ conect_cat,
                    data = data_repro_split) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  scale_colour_viridis_d() +
  scale_fill_viridis_d()
```
### DBH

```{r}
ggplot(survival_data_sc, aes(dbh_sc)) + 
  geom_histogram(bins=20)

data_repro_split <- survival_data_sc %>% 
  mutate(conect_cat = cut(dbh_sc, breaks= 3)) 

ggsurvfit::survfit2(Surv(days, status) ~ conect_cat,
                    data = data_repro_split) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  scale_colour_viridis_d() +
  scale_fill_viridis_d()
```

## Cox

I don't think we can have random effects with the `coxph` function.

```{r}
coxph_mod <- coxph(Surv(days, status) ~ 
                     repro_connectivity_sc +
                     non_repro_connectivity_sc +
                     dbh_sc, 
                   data = survival_data_sc) 

coxph_mod
```

`coxme` package will allow us to add random effects.

```{r}
coxme_mod <- coxme(Surv(days, status) ~ 
                     repro_connectivity_sc +
                     non_repro_connectivity_sc +
                     dbh_sc + (1|tree_id), 
                   data = survival_data_sc) 

coxme_mod
```

One assumption of the Cox proportional hazards regression model is that the hazards are proportional at each point in time throughout follow-up. The `cox.zph()` function from the {survival} package allows us to check this assumption. It results in two main things:

* A hypothesis test of whether the effect of each covariate differs according to time, and a global test of all covariates at once.
  * This is done by testing for an interaction effect between the covariate and log(time)
  * A significant p-value indicates that the proportional hazards assumption is violated
* Plots of the Schoenfeld residuals
  * Deviation from a zero-slope line is evidence that the proportional hazards assumption is violated

```{r}
cox.zph(coxme_mod)

par(mfrow=c(1,3))
cox.zph(coxme_mod) %>% 
  plot() 
```

The proportional hazards assumption is not satisfied for the model overall (p <0.05 for `GLOBAL`).

To relax the PH assumptions we can add an interaction between the predictor and time.
So we are assuming the effect of non-reproductive conspecific density varies linearly with time.. 

```{r}
coxme_mod_2 <- coxme(Surv(days, status) ~ 
                     repro_connectivity_sc +
                     non_repro_connectivity_sc +
                     (non_repro_connectivity_sc:days) +
                     dbh_sc + (1|tree_id), 
                   data = survival_data_sc) 

coxme_mod_2
```

```{r}
cox.zph(coxme_mod_2)

par(mfrow=c(2,3))
cox.zph(coxme_mod_2) %>% 
  plot() 
```

# Bayes

```{r}
bprior <- c(prior(normal(0, 1), class = b))
```


## Cox

```{r warning=FALSE}
options(brms.file_refit = "on_change")

dispersal_mod_cox <-
  brm(data = survival_data_sc,
      family = cox(link = "log", bhaz = NULL),
      days| cens(1-status) ~
        repro_connectivity_sc +
        non_repro_connectivity_sc +
        dbh_sc +
        (1|tree_id),
      prior = bprior,
      iter = 2000,
      warmup = 1000,
      chains = 4,
      cores = 4,
      seed = 9,
      file = (here::here("output", "models", "survival_fit.rds")))

dispersal_mod_cox
```

The coefficients are very similar to our frequentist `coxme_mod` model - a good sign! - but only when using flat priors.

It seems there isn't any way to sample from the posterior for cox models (see issue [here).](https://github.com/paul-buerkner/brms/issues/966)

Might be able to do it following steps [here](https://discourse.mc-stan.org/t/manual-absolute-predictions-from-cox-model-in-brms/27966/8)

```{r}
bhaz <- basehaz(coxph(Surv(days, status) ~ 1, data = survival_data_sc))
```

This gives predictions of survival at 59 days for an average tree:

```{r}
exp(-bhaz$hazard[bhaz$time == 59] * 
      exp(TRUE * as_draws_df(dispersal_mod_cox)$sd_tree_id__Intercept)) |> 
  quantile(probs = c(0.025, 0.5, 0.975))
```

```{r}
tidybayes::tidy_draws(dispersal_mod_cox) %>%
  rename(
    `Diameter at\nbreast height` = b_dbh_sc,
    `Reproductive\nconspecific density` = b_repro_connectivity_sc,
    `Non-reproductive\nconspecific density` = b_non_repro_connectivity_sc
  ) %>%
  select(
    `Reproductive\nconspecific density`,
    `Non-reproductive\nconspecific density`,
    `Diameter at\nbreast height`
  ) %>%
  pivot_longer(cols = everything(), names_to = "parameter") %>%
  ggplot(aes(x = value,
             y = parameter
             )) +
  ggdist::stat_halfeye() +
  geom_vline(xintercept = 0,
             linetype = 2,
             linewidth = 1) +
  labs(x = "Parameter value", y = "") +
  theme(legend.position = "none")
```


## Weibull 

```{r eval=FALSE}
dispersal_mod_weibull <-
  brm(data = survival_data_sc,
      family = weibull(link = "log", link_shape = "log"),
      survey| cens(1-status) ~
        repro_connectivity_sc +
        non_repro_connectivity_sc +
        dbh_sc +
        (1|tree_id),
      prior = bprior,
      iter = 2000,
      warmup = 1000,
      chains = 4,
      cores = 4,
      seed = 9)
```

```{r eval=FALSE}
pp_check(dispersal_mod_weibull, 
         ndraws = 50, 
         type = "km_overlay", 
         status_y = survival_data_sc$status)
```

Notes on Weibull parameters:
For anyone else interested, I believe the following produces the correct survival function for the Weibull distribution, as parameterized in brms. We need parameters lamba and k (the shape parameter estimated in brms). The following R code converts the mean value to lambda:

lambda = exp(mu) / gamma( 1 + 1/k )

Survival at time t is then:

exp( - (t / lambda)^k )

## Log-normal 

```{r eval=FALSE}
dispersal_mod_lognormal <-
  brm(data = survival_data_sc,
      family = lognormal(link = "identity", link_sigma = "log"),
      days| cens(1-status) ~
        repro_connectivity_sc +
        non_repro_connectivity_sc +
        dbh_sc +
        (1|tree_id),
      prior = bprior,
      iter = 2000,
      warmup = 1000,
      chains = 4,
      cores = 4,
      seed = 9)
```

Notes on Lognormal parameters: https://discourse.mc-stan.org/t/understanding-shifted-lognormal-parameters-and-priors/13519/3
plotting the survival curve: https://discourse.mc-stan.org/t/plotting-survival-plots-after-brm/14226

```{r eval=FALSE}
pp_check(dispersal_mod_lognormal, 
         ndraws = 50, 
         type = "km_overlay", 
         status_y = survival_data_sc$status)
```

## Gamma

```{r eval=FALSE}
dispersal_mod_Gamma <-
  brm(data = survival_data_sc,
      family = Gamma(link = "log"),
      survey| cens(1-status) ~
        repro_connectivity_sc +
        non_repro_connectivity_sc +
        dbh_sc +
        (1|tree_id),
      prior = bprior,
      iter = 2000,
      warmup = 1000,
      chains = 4,
      cores = 4,
      seed = 9)

pp_check(dispersal_mod_Gamma)
```

```{r eval=FALSE}
pp_check(dispersal_mod_Gamma, 
         ndraws = 50, 
         type = "km_overlay", 
         status_y = survival_data_sc$status)
```

## Compare models

```{r eval=FALSE}
loo_compare(waic(dispersal_mod_Gamma), 
            waic(dispersal_mod_lognormal), 
            waic(dispersal_mod_weibull), 
            waic(dispersal_mod_cox))
```

# preds for log normal 

```{r eval=FALSE}
survival_data_sc %>%
  modelr::data_grid(
    non_repro_connectivity_sc = modelr::seq_range(non_repro_connectivity_sc, n = 5),
    repro_connectivity_sc = modelr::seq_range(repro_connectivity_sc, n = 5),
    dbh_sc = modelr::seq_range(dbh_sc, n = 5),
    survey = modelr::seq_range(survey, n = 5)
  ) %>%
  tidybayes::add_epred_draws(dispersal_mod_lognormal, ndraws = 500, re_formula = NA,
                             status_y = survival_data_sc$status) -> surv_pred

rstanarm::posterior_survfit(object = dispersal_mod_lognormal) |>
  glimpse()

surv_pred |>
  ggplot(aes(x = repro_connectivity_sc)) +
  ggdist::stat_lineribbon(
    aes(y = .epred )) 

surv_pred |>
  ggplot(aes(x = survey)) +
  ggdist::stat_lineribbon(
    aes(y = .epred )) 
```


# test with simulated data

https://github.com/anddis/brms-weibullPH/tree/main

```{r eval=FALSE}
simdata <- local({
  set.seed(1901)
  N <- 1000
  x <- rbinom(N, 1, 0.5)
  z <- rnorm(N)
  y <- flexsurv::rweibullPH(N, 
                            shape =  1.2, # gamma
                            scale = exp(0 + log(2)*x + log(0.75)*z)) # mu
  cens <- runif(N, 0, 4)
  time <- pmin(y, cens)
  status <- as.numeric(y <= cens)
  data.frame(
    time = time,
    status = status,
    censored = 1 - status,
    x = factor(x),
    z = z
  )
})
```

```{r eval=FALSE}
simdata %>% 
  group_by(time) %>% 
  count(status) %>% 
  pivot_wider(names_from = status, values_from = n) %>% 
  mutate(`0` = replace_na(`0`, 0),
         `1` = replace_na(`1`, 0),)%>% 
  rename(alive = `1`, dead = `0`) %>% 
  group_by(time) %>%
  summarise(total = n()) %>% glimpse
  ggplot(aes(x = time, y = dead/(dead + alive))) +
  geom_point()
```


```{r eval=FALSE}
test_lognormal <-
  brm(data = simdata,
      family = lognormal(link = "identity", link_sigma = "log"),
      time| cens(censored) ~
        x + z ,
      prior = bprior,
      iter = 2000,
      warmup = 1000,
      chains = 4,
      cores = 4,
      seed = 9)
```

```{r eval=FALSE}
pp_check(test_lognormal, 
         ndraws = 50, 
         type = "km_overlay", 
         status_y = simdata$status)
```
