---
title: "Exploring the herbivory data"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
```

```{r}
readRDS(here::here("data", "clean", "connectivity_data.rds")) -> connectivity

read_csv(here::here("data", "raw", "hawthorn_herbivory.csv")) -> herbivory

```
```{r}
herbivory %>%
  dplyr::filter(!grepl("S", sample)) %>%
  mutate(focal_tree = as.integer(tree_id)) %>%
  select(tree_id, focal_tree, sample, date,
         n_leaves_assessed, leaves_w_herbivory) %>%
  inner_join(connectivity, by = c("focal_tree" = "plot")) -> clean

readRDS(here::here("data", "clean", "hawthorn_plots.rds")) %>%
  filter(tree_id == "tree_0") %>%
  mutate(plot = as.numeric(plot)) %>%
  select(plot, dbh) %>%
  left_join(clean, by = c("plot" = "focal_tree"),
             multiple = "all") %>% 
  mutate(is_focal = case_when( floor(tree_id) == tree_id ~ TRUE,
                               .default = FALSE)) -> clean_dbh

```


```{r}
ggplot(filter(clean_dbh, is_focal == TRUE), aes(y = leaves_w_herbivory/n_leaves_assessed, x = connectivity)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
ggplot(filter(clean_dbh, is_focal == TRUE), aes(y = leaves_w_herbivory/n_leaves_assessed, x = dbh)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
clean_dbh %>%
  mutate(connectivity_sc = scale(connectivity),
         dbh_sc = scale(dbh),
         focal_tree = as.factor(tree_id)
  ) -> clean_dbh_sc

lme4::glmer(formula = cbind(leaves_w_herbivory, (n_leaves_assessed - leaves_w_herbivory)) ~ 
     connectivity_sc + dbh_sc + (1|tree_id), 
   data = filter(clean_dbh_sc, is_focal == TRUE), 
   family = binomial(link = "logit")) -> mod

summary(mod)
```

No trend here but looks like there is a lot of variability within a single tree.

## Variability in herbivory

Function to calculate the gini index:

```{r}
get_gini <- function(x, n = rep(1, length(x)) ) {
  x <- rep(x, n)
  n <- length(x)
  x <- sort(x)
  res <- 2 * sum(x * 1:n) / (n*sum(x)) - 1 - (1/n)
  res <- n / (n - 1) * res

  return(res)

}
```

Within tree variability

```{r}
clean_dbh_sc %>%
  filter(is_focal == TRUE) %>% 
  group_by(tree_id) %>%
  mutate(gini = get_gini(x = leaves_w_herbivory/n_leaves_assessed)) -> gini_data

gini_data %>% 
  ggplot(aes(y = gini, x = connectivity)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
lm(gini ~ scale(connectivity), data = gini_data) -> gini_mod

summary(gini_mod)

```

More connected trees have less variability in herbivory within a single tree.
More connectivity -> more herbivory events -> low variability in herbivory

Note that in this analysis I have only used focal trees, 
but we also have data for nearest neighbours and saplings.
To add these trees to the analysis I'd need to calculate connectivity for these trees independently and do some edge effect mitigation.
